### 2014-11-20 doc/thesis/system.tex
Fixing #14
Starting to implement the typing rules in Haskell

@@

Using the rule \textsc{T-IF2}, Typed Lua can type check the following example:
\begin{center}
\begin{tabular}{llll}
\multicolumn{4}{l}{$\mathbf{local} \; o:\String \times \String \cup \Integer \rightarrow \String =$}\\
& \multicolumn{3}{l}{$\mathbf{fun} \; (a:\String, b:\String \cup \Integer):\String$}\\
& & \multicolumn{2}{l}{$\mathbf{if} \; type(b) == ``string" \; \mathbf{then}$}\\
& & & \multicolumn{1}{l}{$\mathbf{return} \; a \; {..} \; b$}\\
& & \multicolumn{2}{l}{$\mathbf{else}$}\\
& & & \multicolumn{1}{l}{$\mathbf{return} \; rep(a, b)_{e}$}\\
\multicolumn{4}{l}{$\mathbf{in} \; o(``foo", 2)_{s}$}
\end{tabular}
\end{center}

@@

Overloading on the return type.
\begin{center}
\begin{tabular}{lll}
\multicolumn{3}{l}{$\mathbf{local} \; q:\pi_{1}, r:\pi_{2} = idiv(1, 2) \; \mathbf{in}$}\\
& \multicolumn{2}{l}{$\mathbf{if} \; type(r) == ``string" \; \mathbf{then}$}\\
& & \multicolumn{1}{l}{$print(``ERROR: " \; .. \; r)_{s}$}\\
& \multicolumn{2}{l}{$\mathbf{else}$}\\
& & \multicolumn{1}{l}{$print(q + r)_{s}$}
\end{tabular}
\end{center}

>>>>

\textit{/* The following example does not type check */}
\begin{center}
\begin{tabular}{ll}
\multicolumn{2}{l}{$\mathbf{local} \; x:\Nil \cup \False = \mathbf{false} \; \mathbf{in}$}\\
& \multicolumn{1}{l}{$\mathbf{local} \; y:\String = x \; \mathbf{or} \; ``Hello"$}
\end{tabular}
\end{center}

\textit{/* Should we define the following rules? */}
\[
\begin{array}{c}
\begin{array}{c}
\mylabel{T-OR4}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \cup \False \subtype t_{1} \;\;\;
       t_{1} \subtype \Nil \cup \False}
      {\env_{1} \vdash e_{1} \; \mathbf{or} \; e_{2}:t_{2}, \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-OR5}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \subtype t_{1} \;\;\;
       \False \subtype t_{1}}
      {\env_{1} \vdash e_{1} \; \mathbf{or} \; e_{2}:filter(filter(t_{1}, \Nil), \False) \cup t_{2}, \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-OR6}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \subtype t_{1} \;\;\;
       \False \not\subtype t_{1}}
      {\env_{1} \vdash e_{1} \; \mathbf{or} \; e_{2}:filter(t_{1}, \Nil) \cup t_{2}, \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-OR7}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \not\subtype t_{1} \;\;\;
       \False \subtype t_{1}}
      {\env_{1} \vdash e_{1} \; \mathbf{or} \; e_{2}:filter(t_{1}, \False) \cup t_{2}, \env_{3}}
\end{array}
\\ \\
\begin{array}{c}
\mylabel{T-OR8}\\
\dfrac{\env_{1} \vdash e_{1}:t_{1}, \env_{2} \;\;\;
       \env_{2} \vdash e_{2}:t_{2}, \env_{3} \;\;\;
       \Nil \not\subtype t_{1} \;\;\;
       \False \not\subtype t_{1}}
      {\env_{1} \vdash e_{1} \; \mathbf{or} \; e_{2}:t_{1} \cup t_{2}, \env_{3}}
\end{array}
\end{array}
\]

\textit{/* Or who writes this kind of code deserves a type error? */}

@@

Using the rule \textsc{T-IF2}, Typed Lua can type check the following example:
\begin{center}
\begin{tabular}{llll}
\multicolumn{4}{l}{$\mathbf{local} \; o:\String \times \String \cup \Integer \rightarrow \String =$}\\
& \multicolumn{3}{l}{$\mathbf{fun} \; (a:\String, b:\String \cup \Integer):\String$}\\
& & \multicolumn{2}{l}{$\mathbf{local} \; r:\String = ``" \; \mathbf{in}$}\\
& & & \multicolumn{1}{l}{$\mathbf{if} \; type(b) == ``string" \; \mathbf{then} \; r = a \;{..}\;b \; \mathbf{else} \; r = rep(a, b)_{e}\; ;$}\\
& & & \multicolumn{1}{l}{$\mathbf{return} \; r$}\\
\multicolumn{4}{l}{$\mathbf{in} \; o(``foo", 2)_{s}$}
\end{tabular}
\end{center}

@@

Overloading on the return type.
\[
\begin{array}{c}
\mylabel{T-IF3}\\
\dfrac{\begin{array}{c}
       \env_{1}(n) = \pi_{i}\\
       closeall(\env_{1}[\pi \mapsto fpt(\env_{1}(\pi), \Nil, i)]) \vdash s_{1}:\env_{2} \\
       closeall(\env_{1}[\pi \mapsto gpt(\env_{1}(\pi), \Nil, i)) \vdash s_{2}:\env_{3}
      \end{array}}
      {\env_{1} \vdash \mathbf{if} \; n \; \mathbf{then} \; s_{1} \; \mathbf{else} \; s_{2}:closeset(\env_{2}, fav(s_{1}) \cup fav(s_{2}))}

\begin{center}
\begin{tabular}{ll}
\multicolumn{2}{l}{$\mathbf{local} \; q:\pi_{1}, r:\pi_{2} = idiv(1, 2)_{m} \; \mathbf{in}$}\\
& \multicolumn{1}{l}{$\mathbf{if} \; q \; \mathbf{then} \; print(q + r)_{s} \; \mathbf{else} \; print(``ERROR: " \; .. \; r)_{s}$}
\end{tabular}
\end{center}

\textit{/* How should we write the typing rule of the local declaration? */}

\textit{/* How this is valid if projection types do not have subtyping rules? */}
